The Array of Long-Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-05-06 albatros_paper.tex
\attention macro added for flagging issues

% Define stuff here
\def\albatros{ALBATROS}
\def\prizm{PRI$^{\rm Z}$M}

@@

\begin{abstract}
\textcolor{red}{\bf text to be overhauled later...} \\
The low frequency radio astronomy has the highest potential in
discovering the history of the Universe, this includes observations
of the first stars and the mapping of dark ages. The Array of Long
Baseline Antennas for Taking Radio Observations from the
Sub-Antarctic (ALBATROS) is a new interferometric array. It consists
of autonomous antenna stations that will map the low-frequency sky
from Marion Island. One autonomous station was deployed in Marion
Island in April 2019. The operating frequency range is
\SIrange{1.2}{81}{\MHz} with baselines of $\approx \SI{20}{\km}$. A
two element pathfinder was deployed in Marion Island in April
2018.
\end{abstract}

@@

\textcolor{red}{\bf Open task for any/all: fix references, make a proper bibtex file}

@@

Very few experiments have measured the radio sky at $\lesssim 30$~MHz.
Comprehensive reviews exist elsewhere (\textcolor{red}{\bf find a good
  reference}), and here, we highlight only a few specific examples.
At the very lowest frequencies, the state of the art among
ground-based measurements dates from the 1950s, when Reber and Ellis
caught brief glimpses of the $2.1$~MHz sky at $\sim
5$\degree\ resolution, using an array of 192~dipoles~\citep{reber,
  article, 1988A&A...195..372W}.  A few space-based missions have also
performed measurements at similarly low frequency ranges; for example,
the Radio Astronomy Explorer-2 operated at \SIrange{0.025}{13}{\MHz}
with $\sim 10$\degree\ resolution at
\SI{4.7}{\MHz}~\citep{1975A&A....40..365A}.  Measurements with
resolution finer than few-degree scales exist only at higher
frequencies.  For example, the Dominion Radio Astrophysical
Observatory surveyed the northern sky at \SI{22}{\MHz} with
\SIrange{1.1}{1.7}{\degree} resolution~\citep{1999A&AS..137....7R},
and most recently, the Owens Valley Long Wavelength Array mapped the
sky with \SI{15}{\arcminute} resolution between 36.528~MHz and
73.152~MHz~\citep{2018AJ....156...32E}.  Although this experimental
list is not comprehensive, it does illustrate the dearth of
information about the $\lesssim 30$~MHz sky and the lack of high
resolution measurements at the lowest frequencies.

\textcolor{red}{\bf More text about the science case, including
  Galactic astrophysics (Jon? Jeff?)}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/marion_map/marion_map_annotated.jpg}
    \caption{Map of Marion Island.  The \albatros\ pathfinder antennas
      are currently installed at the \prizm\ site and at the hydro
      shack.  The black markers denote the eight coastal huts, which
      will be used for future \albatros\ antenna installations.  The
      white markers denote other available infrastructure points that
      will not be used for antennas.  \textcolor{red}{\bf Add new
        synthesized beam figure (Jon)}}
    \label{Fig:marion}
  \end{center}
\end{figure}

@@

The primary requirements that drive the design for a low-frequency
imaging experiment are 1) desired resolution, 2) low RFI, and 3) quiet
ionospheric conditions.  As a benchmark, an interferometer operating
at 30~MHz requires order-of-magnitude baseline lengths of $\sim 1$~km
to achieve a resolution of $\sim 1$\degree.  This length scales
inversely with frequency, and therefore $\sim 10$-km lengths are
required at few-MHz frequencies in order to improve upon the
resolutions achieved to date.  The experiment installation site must
be remote to keep RFI to a minimum, and polar or near-polar latitiudes
generally have lower ionospheric plasma frequency cutoffs.
\textcolor{red}{\bf Add more ionosphere text here, plus IRI simulation
  fig/numbers (Jon)}

Marion Island is a research base that is located in the southern
Indian Ocean at \ang{46;54;45}S, \ang{37;44;37}E and is operated by
the South African National Antarctic Programme.  The island lies
roughly \SI{2000}{\kilo\metre} from the nearest continental landmasses
and has an exceptionally quiet RFI
environment~\citep{2019JAI.....850004P}.  As illustrated in
\autoref{Fig:marion}, Marion has an area of 335~km$^2$ and can
therefore support antenna installations with $>10$-km baseline
lengths.  The main Marion base is located on the northeast side of the
island, and there are eight rest huts along the coast (Cape Davis,
Grey-headed, Kildalkey, Mixed Pickle, Repettos, Rooks, Swartkops,
Watertunnel) and one in the interior (Katedraal) that can serve as
existing infrastructure points for antenna installations.  The planned
\albatros\ installation sites include the coastal huts, but exclude
the main base and Katedraal for RFI and accessibility reasons,
respectively.  \autoref{Fig:marion} also shows the locations of the
\albatros\ pathfinder antennas that are currently installed at the
\prizm\ site and at the hydro shack.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/albatros_2elem/albatros_2elem.pdf}
    \caption{The two-element, directly correlated
      \albatros\ pathfinder installed at the \prizm\ site.  Two
      dual-polarization antennas are separated by roughly 110~m on an
      east--west baseline. Coaxial cables connect the antennas to a
      shipping container that houses the readout electronics and
      serves as the ``command module.'' \textcolor{red}{\bf Add FEE
        and backend sub-pics here? (HCC)}}
    \label{Fig:albatros2}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\linewidth]{Figures/albatros_2elem_schematic/albatros_2elem_schematic.pdf}
    \caption{Schematic for the two-element \albatros\ pathfinder.
      Signals from two dual-polarzation LWA antennas are amplified by
      front-end active baluns (\textcolor{red}{\bf citation for FEE}),
      and 100-m coaxial cables connect the antennas to the back-end
      readout electronics, which are housed in a Faraday cage denoted
      by the dashed box.  Each of the four antenna outputs is passed
      to a second-stage electronics chain consisting of a bias tee,
      amplifier, and high- and low-pass filters.  The signals are
      digitized at 250~Msamp/s by a SNAP board, and the on-board FPGA
      is programmed with firmware to perform a full cross-correlation
      of all four inputs.  A Raspberry Pi controls the SNAP board and
      also saves the computed spectra to an SD card.  The clock signal
      is provided by a Valon frequency synthesizer.  Power to the
      system is provided by a bank of 12~V batteries and several
      regulated voltage outputs.}
    \label{Fig:albatros2_schem}
  \end{center}
\end{figure}

@@

\textcolor{red}{\bf ALL TEXT BELOW HERE NEEDS WORK}

\subsection{Front-end active balun}
\textcolor{red}{\bf Tankiso, Eamon, Jeff?} \\

@@

\subsection{Back-end electronics}
\textcolor{red}{\bf Nivek, HCC?} \\

@@

A Xilinx Kintex\footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html} based Smart Network ADC Processor (SNAP) board (\textcolor{red}{\bf citation for SNAP board?}), 
clocked by a Valon 5007 frequency synthesizer module,
digitizes the RF signals at 250~Msamp/s. It calculates a full
cross-correlation of the four inputs, producing four auto- and six
cross-spectra as outputs, over 2048 channels spanning the frequency range 
\SI{0}{MHz} - \SI{125}{MHz}. 
(\textcolor{red}{\bf details about SNAP board frequency transform?})
The SNAP board sends its output data via ethernet to a Raspberry Pi (\textcolor{red}{\bf version? citation?}), which saves it to a hard drive.

@@

\textcolor{red}{\bf finish merging text
  below into the above paragraph}

@@

\textbf{SNAP Board} - The SNAP board used is specified by Xilinx
Inc. \footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html}. 
% EE it's unclear what this means. The reference is to the Kintex FPGA family, which says nothing about the SNAP board
The
SNAP board samples the signals it receives by 250 Msampl/s using
internal ADCs where the signal of the clock is from the Valon 5007
frerquency synthesizer module. This creates a frequency range between
\SI{0}{MHz} - \SI{125}{MHz} which contains 2048 channels. A SNAP board
interacts with a raspberry pi and that is where the data is saved.
\textcolor{red}{\bf include text about firmware}

@@

\subsection{Software and data acquisition}
\textcolor{red}{\bf Nivek?} \\
Text goes here.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/autonomous.jpg}
    \caption{Single autonomous station pathfinder installed at the
      hydro shack site.  The system is powered by a bank of solar
      panels that are visible in the background.  \textcolor{red}{\bf
        Add command module and electronics glamor shots.}}
    \label{Fig:autonomous}
  \end{center}
\end{figure}

@@

\textcolor{red}{\bf Nivek?} \\

@@

\subsection{Readout electronics}
\textcolor{red}{\bf Nivek, Jon?} \\
Text here about one-bit stuff, pushing baseband through the RPi3,
clock requirements, data rates, etc.

@@

\subsection{Solar power system}
\textcolor{red}{\bf Nivek, Tankiso, Eamon?} \\
The power system of the ALBATROS-EGG uses generators to charge the batteries manually every once in a number of days. From the observation that has been made, the batteries take $\approx$ 2 days to discharge to a point where the system shuts down. Because of the weather conditions in Marion Island, an individual can sometimes be unable to go to the site to charge the batteries, which means that if the weather does not allow, the system can be shut down for a long period. Thus, a solar power system solution was introduced to the ALBATROS so that the system can continously run without having to be recharged manually and often.\\

@@

\textcolor{red}{\bf NEED UPDATED AND NEW PLOTS, NEED VOLUNTEERS FOR THIS:
  \begin{itemize}
    \item{Improved waterfall plots from 2-element pathfinder}
    \item{Plot showing that the solar power works (sort of)}
    \item{Some sort of...data...plot for the autonomous station.  Can
      we say anything about baseband performance?}
  \end{itemize}
}

>>>>

% Define stuff here
\def\albatros{ALBATROS}
\def\prizm{PRI$^{\rm Z}$M}
\newcommand{\attention}[1]{\textcolor{red}{\bf {#1}}}

@@

\begin{abstract}
\attention {text to be overhauled later...} \\
The low frequency radio astronomy has the highest potential in
discovering the history of the Universe, this includes observations
of the first stars and the mapping of dark ages. The Array of Long
Baseline Antennas for Taking Radio Observations from the
Sub-Antarctic (ALBATROS) is a new interferometric array. It consists
of autonomous antenna stations that will map the low-frequency sky
from Marion Island. One autonomous station was deployed in Marion
Island in April 2019. The operating frequency range is
\SIrange{1.2}{81}{\MHz} with baselines of $\approx \SI{20}{\km}$. A
two element pathfinder was deployed in Marion Island in April
2018.
\end{abstract}

@@

\attention{Open task for any/all: fix references, make a proper bibtex file}

@@

Very few experiments have measured the radio sky at $\lesssim 30$~MHz.
Comprehensive reviews exist elsewhere (\attention{find a good
  reference}), and here, we highlight only a few specific examples.
At the very lowest frequencies, the state of the art among
ground-based measurements dates from the 1950s, when Reber and Ellis
caught brief glimpses of the $2.1$~MHz sky at $\sim
5$\degree\ resolution, using an array of 192~dipoles~\citep{reber,
  article, 1988A&A...195..372W}.  A few space-based missions have also
performed measurements at similarly low frequency ranges; for example,
the Radio Astronomy Explorer-2 operated at \SIrange{0.025}{13}{\MHz}
with $\sim 10$\degree\ resolution at
\SI{4.7}{\MHz}~\citep{1975A&A....40..365A}.  Measurements with
resolution finer than few-degree scales exist only at higher
frequencies.  For example, the Dominion Radio Astrophysical
Observatory surveyed the northern sky at \SI{22}{\MHz} with
\SIrange{1.1}{1.7}{\degree} resolution~\citep{1999A&AS..137....7R},
and most recently, the Owens Valley Long Wavelength Array mapped the
sky with \SI{15}{\arcminute} resolution between 36.528~MHz and
73.152~MHz~\citep{2018AJ....156...32E}.  Although this experimental
list is not comprehensive, it does illustrate the dearth of
information about the $\lesssim 30$~MHz sky and the lack of high
resolution measurements at the lowest frequencies.

\attention{More text about the science case, including
  Galactic astrophysics (Jon? Jeff?)}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/marion_map/marion_map_annotated.jpg}
    \caption{Map of Marion Island.  The \albatros\ pathfinder antennas
      are currently installed at the \prizm\ site and at the hydro
      shack.  The black markers denote the eight coastal huts, which
      will be used for future \albatros\ antenna installations.  The
      white markers denote other available infrastructure points that
      will not be used for antennas.  \attention{Add new
        synthesized beam figure (Jon)}}
    \label{Fig:marion}
  \end{center}
\end{figure}

@@

The primary requirements that drive the design for a low-frequency
imaging experiment are 1) desired resolution, 2) low RFI, and 3) quiet
ionospheric conditions.  As a benchmark, an interferometer operating
at 30~MHz requires order-of-magnitude baseline lengths of $\sim 1$~km
to achieve a resolution of $\sim 1$\degree.  This length scales
inversely with frequency, and therefore $\sim 10$-km lengths are
required at few-MHz frequencies in order to improve upon the
resolutions achieved to date.  The experiment installation site must
be remote to keep RFI to a minimum, and polar or near-polar latitiudes
generally have lower ionospheric plasma frequency cutoffs.
\attention{Add more ionosphere text here, plus IRI simulation
  fig/numbers (Jon)}

Marion Island is a research base that is located in the southern
Indian Ocean at \ang{46;54;45}S, \ang{37;44;37}E and is operated by
the South African National Antarctic Programme.  The island lies
roughly \SI{2000}{\kilo\metre} from the nearest continental landmasses
and has an exceptionally quiet RFI
environment~\citep{2019JAI.....850004P}.  As illustrated in
\autoref{Fig:marion}, \attention{this points to the map, not the beam pattern} Marion has an area of 335~km$^2$ and can
therefore support antenna installations with $>10$-km baseline
lengths.  The main Marion base is located on the northeast side of the
island, and there are eight rest huts along the coast (Cape Davis,
Grey-headed, Kildalkey, Mixed Pickle, Repettos, Rooks, Swartkops,
Watertunnel) and one in the interior (Katedraal) that can serve as
existing infrastructure points for antenna installations.  The planned
\albatros\ installation sites include the coastal huts, but exclude
the main base and Katedraal for RFI and accessibility reasons,
respectively.  \autoref{Fig:marion} also shows the locations of the
\albatros\ pathfinder antennas that are currently installed at the
\prizm\ site and at the hydro shack.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/albatros_2elem/albatros_2elem.pdf}
    \caption{The two-element, directly correlated
      \albatros\ pathfinder installed at the \prizm\ site.  Two
      dual-polarization antennas are separated by roughly 110~m on an
      east--west baseline. Coaxial cables connect the antennas to a
      shipping container that houses the readout electronics and
      serves as the ``command module.'' \attention{Add FEE
        and backend sub-pics here? (HCC)}}
    \label{Fig:albatros2}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\linewidth]{Figures/albatros_2elem_schematic/albatros_2elem_schematic.pdf}
    \caption{Schematic for the two-element \albatros\ pathfinder.
      Signals from two dual-polarzation LWA antennas are amplified by
      front-end active baluns (\attention{citation for FEE}),
      and 100-m coaxial cables connect the antennas to the back-end
      readout electronics, which are housed in a Faraday cage denoted
      by the dashed box.  Each of the four antenna outputs is passed
      to a second-stage electronics chain consisting of a bias tee,
      amplifier, and high- and low-pass filters.  The signals are
      digitized at 250~Msamp/s by a SNAP board, and the on-board FPGA
      is programmed with firmware to perform a full cross-correlation
      of all four inputs.  A Raspberry Pi controls the SNAP board and
      also saves the computed spectra to an SD card.  The clock signal
      is provided by a Valon frequency synthesizer.  Power to the
      system is provided by a bank of 12~V batteries and several
      regulated voltage outputs.}
    \label{Fig:albatros2_schem}
  \end{center}
\end{figure}

@@

\attention{ALL TEXT BELOW HERE NEEDS WORK}

\subsection{Front-end active balun}
\attention{Tankiso, Eamon, Jeff?} \\

@@

\subsection{Back-end electronics}
\attention{Nivek, HCC?} \\

@@

A Xilinx Kintex\footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html} based Smart Network ADC Processor (SNAP) board (\attention{citation for SNAP board?}), 
clocked by a Valon 5007 frequency synthesizer module,
digitizes the RF signals at 250~Msamp/s. It calculates a full
cross-correlation of the four inputs, producing four auto- and six
cross-spectra as outputs, over 2048 channels spanning the frequency range 
\SI{0}{MHz} - \SI{125}{MHz}. 
(\attention{details about SNAP board frequency transform?})
The SNAP board sends its output data via ethernet to a Raspberry Pi (\attention{version? citation?}), which saves it to a hard drive.

@@

\attention{finish merging text
  below into the above paragraph}

@@

\textbf{SNAP Board} - The SNAP board used is specified by Xilinx
Inc. \footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html}. 
% EE it's unclear what this means. The reference is to the Kintex FPGA family, which says nothing about the SNAP board
The
SNAP board samples the signals it receives by 250 Msampl/s using
internal ADCs where the signal of the clock is from the Valon 5007
frerquency synthesizer module. This creates a frequency range between
\SI{0}{MHz} - \SI{125}{MHz} which contains 2048 channels. A SNAP board
interacts with a raspberry pi and that is where the data is saved.
\attention{include text about firmware}

@@

\subsection{Software and data acquisition}
\attention{Nivek?} \\
Text goes here.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/autonomous.jpg}
    \caption{Single autonomous station pathfinder installed at the
      hydro shack site.  The system is powered by a bank of solar
      panels that are visible in the background.  \attention{
        Add command module and electronics glamor shots.}}
    \label{Fig:autonomous}
  \end{center}
\end{figure}

@@

\attention{Nivek?} \\

@@

\subsection{Readout electronics}
\attention{Nivek, Jon?} \\
Text here about one-bit stuff, pushing baseband through the RPi3,
clock requirements, data rates, etc.

@@

\subsection{Solar power system}
\attention{Nivek, Tankiso, Eamon?} \\
The power system of the ALBATROS-EGG uses generators to charge the batteries manually every once in a number of days. From the observation that has been made, the batteries take $\approx$ 2 days to discharge to a point where the system shuts down. Because of the weather conditions in Marion Island, an individual can sometimes be unable to go to the site to charge the batteries, which means that if the weather does not allow, the system can be shut down for a long period. Thus, a solar power system solution was introduced to the ALBATROS so that the system can continously run without having to be recharged manually and often.\\

@@

\attention{NEED UPDATED AND NEW PLOTS, NEED VOLUNTEERS FOR THIS:
  \begin{itemize}
    \item{Improved waterfall plots from 2-element pathfinder}
    \item{Plot showing that the solar power works (sort of)}
    \item{Some sort of...data...plot for the autonomous station.  Can
      we say anything about baseband performance?}
  \end{itemize}
}
